---
title: "phyE mutant analysis"
author: "Julin Maloof"
date: "August 10, 2015"
output: html_document
---

## Load the data and libraries

```{r}
library(ggplot2)
library(lmerTest)
library(lme4)
data <- read.csv("phyintpetbothreps.csv")
head(data)
summary(data)
data <- data[,!grepl("X",colnames(data))] # get rid of time stamp columns
summary(data)
```

## Coefficient of variation for each organ
I am curious as to which data have the least "noise" or residual variation.  One way to do this is to calculate the coefficient of variation.  CV is standard dev / mean.  I am also curious, for the internodes, what combining them looks like.

```{r}
data <- within(data,{
  epi.int1 <- epi+int1
  epi.int123 <- epi+int1+int2+int3
})
with(data,{
  sapply(names(data)[7:16],function(x) {
    means <- tapply(get(x),list(genotype,treatment,rep),mean,na.rm=T)
    sdevs <- tapply(get(x),list(genotype,treatment,rep),sd,na.rm=T)
    mean(sdevs/means,na.rm=T)
  })
})
```

Conclusion: earlier organs are better.  Combining epi+int1 is also good; combining all shoot organs may be reasonable.

## Modelling for individual organs

Use a mixed effect model to take into account random variation, like rep and flat.  Need to make a new variable for flat to make sure we aren't confused about which flat came from which experiment.




